HERMES : Hybrid ERror-corrector Model with inclusion of External Signals for nonstationary fashion time series

### 2022-01-19 hermes_paper/hermes_ieee.tex
Minor corrections in the introduction

\author{Etienne David\thanks{Etienne David: Heuritech, France.}, Jean Bellot\thanks{Jean Bellot: Heuritech, France.} and Sylvain Le Corff\thanks{Sylvain Le Corff:  Samovar, T\'el\'ecom SudParis, d\'epartement CITI, TIPIC, Institut Polytechnique de Paris, France.}}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell was with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
Parametric generative models allow to provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \cite{box2015time} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

In this paper, we consider a  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ weekly sequences  of fashion trends on social media  from 01-01-2015 to 01-01-2019. This dataset has very appealing properties:  all time series have the same length, same seasonality, no missing value and there is no sparse time series even for niche trends. The originality of our dataset comes from the fact that additional external weak signals can also be introduced. With our fashion expertise, we detected several groups of highly influential fashion users. Analyzing their specific behaviours on social media, we associate with each time series an external weak signal representing the same fashion trends on a sub-category of users. They are called weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at designing a model able to deal with such a large dataset, leveraging  complex external weak signals and finally providing the most accurate forecasts.

>>>>

\author{Etienne David\thanks{Etienne David: Heuritech, France.}, Jean Bellot\thanks{Jean Bellot: Heuritech, France.} and Sylvain Le Corff\thanks{Sylvain Le Corff:  Samovar, T\'el\'ecom SudParis, d\'epartement CITI, Institut Polytechnique de Paris, France.}}
%        and~Jane~Doe,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
%\thanks{M. Shell was with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised August 26, 2015.}}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \cite{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \cite{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \cite{box2015time} are for instance very popular parametric generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom have a complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \cite{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \cite{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Although these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

In this paper, we consider a  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ weekly sequences  of fashion trends on social media  from 01-01-2015 to 01-01-2019. This dataset has very appealing properties:  all time series have the same length, no missing value and there is no sparse time series even for niche trends. The originality of our dataset comes from the fact that additional external weak signals can also be introduced. With our fashion expertise, we detected several groups of highly influential fashion users. Analyzing their specific behaviours on social media, we associate with each time series an external weak signal representing the same fashion trends on a sub-category of users. They are called weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at designing a model able to deal with such a large dataset, leveraging  complex external weak signals and finally providing the most accurate forecasts.

### 2022-01-19 hermes_paper/hermes_paper.tex
Minor corrections in the introduction

\author{Etienne DAVID \\
Heuritech \\
Paris \\
\texttt{etienne.david@heuritech.com} \\
\And
Sylvain Le Corff \\
Samovar, T\'el\'ecom SudParis, d\'epartement CITI, TIPIC \\
Institut Polytechnique de Paris, Palaiseau \\
\texttt{sylvain.le$\_$corff@telecom-sudparis.eu} \\
\And 
Jean Bellot \\
Heuritech \\
Paris \\
\texttt{jean.bellot@heuritech.com} \\
}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \citep{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
 Parametric generative models allow to provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \citep{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \citep{box2015time} are for instance very popular parametric linear generative models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom present a highly complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
 In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \citep{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \citep{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Altough these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

In this paper, we consider  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ sequences representing the apparition of fashion trends on social media per week from 01-01-2015 to 01-01-2019. Our fashion time series has very appealing properties: they have all the same length, same seasonality, no missing value and the absence of sparse time series even for niche trends. The originality of our dataset come from the fact that additional external weak signals can be introduced. With our fashion expertise, we detected several groups of fashion users with a high influence power. Analysing their specific behaviours on social medias, we add to each time series, 4 external weak signals representing the same fashion trends but on a sub-category of users. We call them weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at designing a model able to deal with our large fashion dataset, leverage our complex external weak signals and finally provide the most accurate forecasts.

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining deep neural network (DNN) architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citep{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most striking results was proposed in  \citep{smyl2020hybrid} in the context of the M4 forecasting competition \citep{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model is introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network is trained on the entire dataset to correct errors of the previous exponential smoothing models. %With this process, the hybrid approach succeeds at mixing strengths of both model families: robustness of the time series specific parametric models and the power of neural network at learn complex dynamics on large dataset. 
%However, the hybrid model has not been designed to deal with external signals even if it include a neural network.

>>>>

\author{Etienne DAVID \\
Heuritech \\
Paris \\
\texttt{etienne.david@heuritech.com} \\
\And
Sylvain Le Corff \\
Samovar, T\'el\'ecom SudParis, d\'epartement CITI, \\
Institut Polytechnique de Paris, Palaiseau \\
\texttt{sylvain.le$\_$corff@telecom-sudparis.eu} \\
\And 
Jean Bellot \\
Heuritech \\
Paris \\
\texttt{jean.bellot@heuritech.com} \\
}

@@

Multivariate time series forecasting is a widespread statistical problem with  many applications, see for instance \citep{sarkka2013bayesian, douc2014nonlinear, zucchini2017hidden} and the numerous references therein.
 %Due to the diversity of applications and use cases, a multitude of models have been proposed. 
 Parametric generative models provide explainable predictions with statistical guarantees based on a precise modeling of the predictive distributions of new data based on a record of past observations. %The parameters of these models are usually estimated using a sequence of observations of the target time series. 
Calibrating these models, for instance using maximum likelihood inference, often requires a fair amount of tuning to design a time series-specific model to provide  accurate forecasts and sharp confidence intervals.  Depending on the use case, statistical properties of the signal and the available data, many families of models have been proposed for time series.  The exponential smoothing model \citep{RePEc:inm:oropre:v:9:y:1961:i:5:p:673-685}, the Trigonometric Box-Cox transform, ARMA errors, Trend, and Seasonal components model (TBATS) \cite{doi:10.1198/jasa.2011.tm09771}, or the ARIMA with the Box-Jenkins approach \citep{box2015time} are for instance very popular parametric  models.  Hidden Markov models (HMM) are also widespread and presuppose that available observations are defined using missing data describing the dynamical system. This hidden state is assumed to be a Markov chain such that at each time step the received observation is a random function of the corresponding latent data.  Although hidden states are modeled as a Markov chain, the observations arising therefrom present a highly complex statistical structure. %Introducing latent hidden states, they assume that the predictive distribution is not unique and constant in time but multiple and lead by the values of the hidden states changing during time. 
 In various applications where signals exhibit non-stationarities such as trends and seasonality, classical HMM are not adapted. However, \citep{touron2017modeling}  recently proposed seasonal HMM, assuming that transition probabilities between the states, as well as the emission distributions, are not constant in time but evolve in a periodic manner. Strong consistency results were established in \citep{touron2019consistency} and Expectation Maximization based numerical experiments were proposed.
Altough these works provide promising results, HMM are computationally expensive to train and are not yet well studied for seasonal  sequences with thousands of components.

@@

In this paper, we consider  new time series forecasting application referred to as {\em fashion trends prediction}. Based on a cutting-edge image recognition technology, we built the first fashion dataset containing \numberts\ sequences representing the apparition of fashion trends on social media per week from 01-01-2015 to 01-01-2019. Our fashion time series has very appealing properties: they have all the same length, no missing value and the absence of sparse time series even for niche trends. The originality of our dataset come from the fact that additional external weak signals can be introduced. With our fashion expertise, we detected several groups of fashion users with a high influence power. Analysing their specific behaviours on social medias, we add to each time series, 4 external weak signals representing the same fashion trends but on a sub-category of users. We call them weak signals because they are often alerts or events that are too sparse, or too incomplete to allow on their own an accurate estimation of their impact on the prediction of the target signal. With this totally new application, we aim at designing a model able to deal with our large fashion dataset, leverage our complex external weak signals and finally provide the most accurate forecasts.

Recurrent neural networks are appealing to tackle our forecasting problem due to their capability of leveraging external data.  Recently, hybrid models combining deep neural network (DNN) architectures with widespread statistical models to deal with seasonality and trends have been proposed, see for instance  \citep{zhang2003time,jianwei2019novel,bandara2020lstm}. The approach providing the most impressive results was proposed in  \citep{smyl2020hybrid} in the context of the M4 forecasting competition \citep{makridakis2020m4}.  Given a large dataset, a per-time-series multiplicative exponential smoothing model is introduced to estimate simple but fundamental components for each time series and compute a first prediction. Then a global recurrent neural network is trained on the entire dataset to correct errors of the previous exponential smoothing models. %With this process, the hybrid approach succeeds at mixing strengths of both model families: robustness of the time series specific parametric models and the power of neural network at learn complex dynamics on large dataset. 
%However, the hybrid model has not been designed to deal with external signals even if it include a neural network.
